### 2020-09-07 old-versions/dac2020-hardwareKEM/intro.tex
first commit

As in the past case for standardizing AES and SHA-3, the parameters which will be used for selection include the security of the algorithm and its efficiency when implemented in hardware and software. NIST have also stated that algorithms which can be made robust against physical attacks in an effective and efficient way will be preferred \cite{nistsca}. Thus, it is important, during the scrutiny of the candidates, to explore the potential of implementing these algorithms on a variety of platforms, and to assess the overhead of adding countermeasures against physical attacks.

To this end, this paper concentrates on FrodoKEM, a key encapsulation algorithm submitted to NIST as a potential post-quantum standard. FrodoKEM is a conservative candidate due to its hardness being based on standard lattices, as opposed to Ring-LWE or Module-LWE, as such it has had limited practical evaluations. Thus, we explore the possibility to efficiently implementing it in hardware and we estimate the overhead of protection against power analysis attacks using first-order masking. To maximize the throughput, we rely on a parallelised implementation of the matrix multiplication. We also aim to have a relatively low area consumption on the FPGA. To be parallelised, however, the matrix multiplication requires the use of a smaller and more performant pseudo-random number generator. We propose to achieve the performance required for the PRNG by using Trivium, an international standard under ISO/IEC 29192-3 \cite{ISO} and selected as part of the eSTREAM project. We utilize this instead of AES or SHAKE, as per the FrodoKEM specifications. We do this as a design exploration study, but, all the security arguments from the specifications still apply.

>>>>

As in the past case for standardizing AES and SHA-3, the parameters which will be used for selection include the security of the algorithm and its efficiency when implemented in hardware and software. NIST have also stated that algorithms which can be made robust against physical attacks in an effective and efficient way will be preferred \cite{nistsca}. Thus, it is important, during the scrutiny of the candidates, to explore the potential of implementing these algorithms on a variety of platforms, and to assess the overhead of adding countermeasures.

To this end, this paper concentrates on FrodoKEM, a key encapsulation mechanism submitted to NIST as a potential post-quantum standard. FrodoKEM is a conservative candidate due to its hardness being based on standard lattices, as opposed to Ring/Module-LWE, thus having limited practical evaluations. Thus, we explore the possibility to efficiently implementing it in hardware and estimate the overhead of protection against power analysis attacks using first-order masking. To maximize the throughput, we rely on a parallelised implementations of the matrix multiplication. We also aim to have a relatively low FPGA area consumption. To be parallelised, however, the matrix multiplication requires the use of a smaller and more performant pseudo-random number generator. We propose to achieve the performance required for the PRNG by using Trivium, an international standard under ISO/IEC 29192-3 \cite{ISO} and selected as part of the eSTREAM project. We utilize this instead of AES or SHAKE, as per the FrodoKEM specifications. We do this as a design exploration study, but, all the security arguments from the specifications still apply.
